### 2020-10-19 manuscript.Rmd
Initialize manuscript styling

---
documentclass: jss
classoption: nojss
citation_package: natbib
bibliography: references.bib
author:
  - name: Devin Incerti
    affiliation: Genentech
    # use this syntax to add text on several lines
    address: |
      | South San Francisco
    email: \email{incerti.devin@gene.com}
  - name: Jeroen Jansen
    affiliation: |
      | University of California, San Francisco
    address: |
      | San Francisco
    email: \email{jeroen.jansen@ucsf.edu}
title:
  plain:     "hesim: Health Economic Simulation Modeling and Decision Analysis"
  formatted: "\\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis"
abstract: >
  The abstract of the article.
keywords:
  formatted: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, "\\proglang{R}"]
  plain: [health economic evaluation, cost-effectiveness analysis, simulation, modeling, R]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
knitr::opts_chunk$set(eval = TRUE)
```

# Introduction